The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area

### 2024-03-27 MS_ToSubmit/RevisedVersion_1.Rmd
Revised version 1

---
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: yes
  word_document: default
bibliography: AMPBancoBurdwood.bib
csl: "elsevier-harvard2.csl"
documentclass: article
geometry:
- vmargin=1in
- hmargin=1in
header-includes:
- \usepackage{amsmath}
- \usepackage{pdflscape,booktabs}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage[running]{lineno}
- \linenumbers
---

@@

The total area of the world ocean designated under marine protection has increased recently. Most Marine Protected Areas (MPAs) target vulnerable, keystone, charismatic and or endemic species. In the sub-Antarctic, ocean protection is associated to oceanic islands, except for MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, ~53º–55ºS and ~56º–62ºW), which are associated to a submarine plateau and its southern adjacent deep slope, respectively. Here, we present the first analysis of the network of predator-prey interactions for the MPA N-BB. We applied a topological network approach to characterise the complexity and structure of the food web, and identify the species' role in such a framework. The MPA N-BB food web consisted of 1788 interactions and 379 species, with a connectance of 0.01. Almost half of the consumers were omnivores or feeding at more than one trophic level (0.48), and the network displayed a small-world pattern (short path length, high clustering of compartments). This network pattern suggests that the ecosystem might be vulnerable to perturbations targeting highly connected species, although some properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Several species arose as important in terms of different aspects of trophic structure and functioning, and response to perturbations. Generalist species, mainly fishes, play a crucial role in the ecosystem's benthopelagic coupling and should be considered as relevant energy transfers for the ecosystem. We argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for securing the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

@@

Identifying the main species involved in the maintenance of ecosystem services and health as well as for management and conservation is essential. In this regard, @Bergagna2020 have identified the principal ecosystem services that are provided by species inhabiting the MPA N-BB; for instance, nursery and feeding areas for commercially important fishes (e.g. Patagonian grenadier *Macruronus magellanicus*, Patagonian toothfish *Dissostichus eleginoides*), water purification by filter feeders (e.g. sponges), among others. These species are part of a complex system in terms of biodiversity. There is robust knowledge on the complexity considering the richness of the benthic and plankton communities in the MPA N-BB ecosystem [@Schejter2016; @Guinder2020; @Schejter2020; @AdministraciondeParquesNacionales2022]. Overall, 811 benthic and plankton species have been identified in the MPA N-BB ecosystem, where 349 were reported for the first time in the area in recent years [@AdministraciondeParquesNacionales2022]. Recently, the structure of the southwestern South Atlantic Ocean has been hypothesised to be under a 'wasp-waist' control, meaning that the structure and dynamics of the ecosystem are regulated primarily by mid-trophic level species (e.g., fishes, crustaceans) [@Padovani2012; @Saporiti2015; @Riccialdelli2020]. In particular, the ecosystem of the MPA N-BB shows a more pronounced 'wasp-waist' structure, meaning a shorter average food chain and a greater trophic overlap and redundancy, than other sub-Antarctic areas such as the continental shelf off Tierra del Fuego. It is the high abundance of few prey species at this mid-trophic level of the food web which imposes a certain feeding dependence to high-trophic level predators, shortens food-web length and increases the trophic overlap and redundancy of the overall food web. The Fuegian sprat *Sprattus fuegensis* and longtail southern cod Patagonotothen ramsayi are considered the most plausible 'wasp-waist' species, since they are species with high regional abundances [@Madirolas2000; @Arkhipkin2013], occupy mid-trophic levels [@Riccialdelli2010; @Riccialdelli2020], many predators feed on them [@Riccialdelli2020], and the population dynamics of these species appear to depend on the environment (e.g. climatic variations) [@Diez2018]. These recent efforts have increased our understanding of the biological communities and their complexity, but an understanding of species interactions’ complexity and structure is lacking. Such understanding will provide a more complete picture of the ecosystem, generate a baseline to comprehend climate change effects for the region [@Franco2020] and serve as an input to decision makers on conservation actions [@AdministraciondeParquesNacionales2022].

This aspect can be tackled by analysing one of the most frequent relationships between species: the predator-prey interaction [@Bascompte2009]. The sum of these interactions in a particular region is referred to as a food web, representing the roadmap for matter and energy flow in an ecosystem. In recent years, topological network approaches have been successfully applied to study complex high-latitude marine ecosystems, improving our knowledge on structure, functioning, and response to environmental/anthropogenic changes [@Kortsch2015; @Cordone2018; @Funes2022; @Marina2023]. Among anthropogenic threats, it is worth mentioning that contaminants like mercury and microplastics have been recently reported as important threats to the MPA N-BB region [@Cossi2021; @DiMauro2022; @Fioramonti2022]; also fishing vessels are allowed to operate in the western section of the MPA N-BB (i.e. Marine National Reserve category), altering the stocks of commercially important fish species [@Martinez2021; @AdministraciondeParquesNacionales2022]. Moreover, there is a potential hazard related to the effects of offshore activities (exploration and exploitation of gas and oil) to the west of the MPA N-BB [@AdministraciondeParquesNacionales2022].

@@

where $E_i$ is the effective number of interactions between $K_i$ nearest-neighbour nodes of node $i$ and the maximal possible number of such interactions [@Newman2003].
To test whether the food web presented the small-world pattern, we compared the empirical values of $CPL$ and $CC$ with those resulting from 1000 randomly generated networks with the same size ($S$) and number of interactions ($L$). Random webs were created using the Erdös-Rényi model, where links are added to the complete set of nodes ($S$) and chosen uniformly randomly from the set of all possible links [@Erdos1959]. The rigurosity of our method lies in the use of confidence intervals (CI 99%) for the empirical-random comparison of the $CPL$ and $CC$ properties. If the empirical value is positioned within or to the left (= lower than) the CI 99% of the random CPL, and to the right (= higher than) the CI 99% of the CC, then the food web is considered to present the SW topology [@Marina2018a].

@@

The species-level analysis aims to describe the species' role in the food web. For this, we considered the following properties: betweenness $Btw$, closeness $Cl$, trophic similarity $TS$, topological role $TR$, and trophic level $TL$ (Table 1). Topological roles refer to the fact that food webs tend to naturally organize in non-random, modular patterns, where modules are defined as a group of species that interact more frequently among themselves than with species that are not members of the module [@Guimera2005]. This means that groups of prey and predators interact more strongly with each other, than with species belonging to other groups. Modularity measures how strongly these subgroups of species, called modules, interact with each other compared to species from other modules. Species can play different roles in this respect, according to the pattern of interactions within their own module and/or across modules. We computed the topological role for each species, classified as module hub, species with a relatively high number of interactions, but most within its own module; module specialist, species with relatively few interactions and most within its own module; module connector, species with relatively few interactions mainly between modules; and network connector, species with high connectivity between and within modules [@Guimera2005]. Refer to the Supplementary Material for the equations used to calculate the species-level properties.

@@

The degree distribution of the food web showed an asymmetric frequency in the number of interactions, where most of the species had a relatively low number of interactions and few species concentrated most of them (Figure 3A). The distribution of prey among predators showed that most consumers fed on a low number of prey whereas few had multiple prey (Figure 3B). The top-five predators in number of prey were: yellowfin notothen *Patagonotothen guntheri* (Notothenioid fish, 50 prey), rock cod *Patagonotothen ramsayi* (Notothenioid fish, 49 prey), broad nose skate *Bathyraja brachyurops* (Chondrichthyan, 33 prey), Patagonian toothfish *Dissostichus eleginoides* (Notothenioid fish, 30 prey), and graytail skate *Bathyraja griseocauda* (Chondrichthyan, 28 prey). Following the same distribution pattern, few prey presented multiple predators (Figure 3C). The top-five prey (or food sources) in number of predators were: Detritus (Non-living, 153 predators), the three categories of Diatoms considered (benthic, centric and pennate, 72.5 predators on average), and species of the genus *Euphausia* (Zooplankton, 46 predators). Finally, aggregating the interactions by functional group (Figure 3D), most interactions were concentrated in a few species. The most evident species were: *Doryteuthis gahi* (Cephalopoda), *Grimothea (=Munida) gregaria* (Decapoda), *Patagonotothen ramsayi*, *Patagonotothen guntheri* and *Dissostichus eleginoides* (bentho-pelagic fish), *Sprattus fuegensis* and *Micromesistius australis* (pelagic fish), and species of *Euphausia* and *Themisto gaudichaudii* (Zooplankton).
Overall, there is an evident asymmetry in the distribution of interactions among species at different levels (i.e. considering the entire food web, gathered by prey, predator and functional group) in the MPA N-BB food web. A list of the distribution of interactions per species is presented in Supplementary Material (Table S3).

@@

Food web connectance is a feature that resumes the complexity of the network, but more importantly, it is an emergent property of pairwise species interactions [@Poisot2014]. It contains information regarding how interactions within an ecological network are distributed and predicts reasonably well key dynamical properties of ecological networks [@Dunne2002a]. Complex marine food webs (i.e. with more than 25 trophic species) show connectance values ranging from 0.01 - 0.27 [@Marina2018a]. In particular, food webs from high-latitude regions tend to exhibit a connectance closer to the minimum (between 0.01 and 0.05) [@Santana2013; @Kortsch2015; @Rodriguez2022]. The connectance of the food web of the MPA N-BB (0.01) is one of the lowest reported so far for these regions; in particular, it appears to be much lower than that of Beagle Channel (0.05), an adjacent coastal area [@Rodriguez2022]. Whether food webs display a low or a high connectance helps to better comprehend ecosystem's synthetic properties like robustness. Empirical research indicates that ecological networks with high connectance are resilient to external perturbations, including the introduction of new species like invasive ones [@Smith-Ramesh2017], as well as the removal of species due to local extinction [@Dunne2002; @Montoya2003].

@@

Closeness and betweenness are defined as "mesoscale" properties because they consider direct and indirect interactions, therefore describing the focal species’ ability to influence the rest of the species of the food web [@Freeman1978; @Lai2012].
Closeness quantifies how many steps away species $i$ is from all other species in the food web, and is proportional to how rapidly the indirect effects of the focal species can spread to other species in the network [@Scotti2010].
Considering the relationship between this property and TL in the food web of the MPA N-BB, it is worth to highlight that species with the highest closeness values are low-TL consumers (TL = 2 - 3): species of the zooplankton community, *Calanus* and *Euphausia*, *Zygochlamys patagonica* (Bivalvia), and Brachiopoda. Any perturbation affecting these species, such as the recently confirmed contaminants mercury [@Fioramonti2022] and microplastics [@Cossi2021; @DiMauro2022], and the warming trend reported for the surface and subsurface waters in the Southwest Atlantic Ocean [@Franco2020a; @Galvan2022a], should be of concern since it might reach many other species in the food web.
Otherwise, betweenness measures the number of shortest paths between species, providing information on the importance of species as “bridges” for energy transfer: a species with high betweenness takes part in more food chains and therefore affects more energy flows [@Scotti2010]. We have identified the longtail southern cod *Patagonotothen ramsayi* as the most important species in this sense. Moreover, in light of our analysis, other species like the Patagonian toothfish *Dissostichus eleginoides* (juveniles), the Patagonian cod *Salilota australis*, the yellowfin notothenioid *Patagonotothen guntheri*, and the Patagonian longfin squid *Doryteuthis gahi* should be considered as relevant in the energy transfer in the ecosystem. All these species have a mid-trophic position in the food web, supporting the 'wasp-waist' control hypothesis for the MPA N-BB [@Riccialdelli2020].

Ecosystems with a pronounced 'wasp-waist' structure are suggested to present a high trophic redundancy, since many species would show similar trophic habits [@Cury2000]. Such trophic redundancy is expected to be greater for species in TLs above the waist (mid-TL) that depend on a very few species at the waist. However, our results show an exponential-like negative relationship between trophic similarity (or redundancy) and trophic level, indicating functional similarity at low and mid-TL species compared to higher TL species for the MPA N-BB food web.
This contradiction might be reflecting the potential discrepancy that could arise between qualitative (this study) and quantitative food web studies. Here it is important to note that our analyses considered all potential predator-prey interactions in the food web giving equal importance to them. Although this could be a disadvantage when inferring on energy flows, one of the main advantages of qualitative food web studies is that it considers the diversity of the communities inhabiting a particular study area, therefore providing a more representative picture of all components (species) acting on the ecosystem. In this sense, our study highlights the uniqueness in terms of the trophic role of high-TL predators. Not only the expected pelagic animals such as marine mammals and seabirds arise as relevant, but also demersal vertebrate (chondrichthyans *Bathyraja macloviana* and *Squalus acanthias*) and benthic invertebrate species (echinoderms *Diplopteraster clarki* and *Pteraster* sp.) are noteworthy. The role that such species play in the MPA N-BB ecosystem is unique and perturbations on them might result in unprecedented changes at the trophic structure and functioning level. In this regard, we should mention the potential threat of the fisheries operating in the western section of the MPA N-BB, where this activity is allowed and mostly focuses on the Patagonian toothfish *Dissostichus eleginoides* and the southern blue whiting *Micromesistius australis* [@Martinez2015]. Although the fishing effort is concentrated outside the limits of the MPA N-BB, the impact on the MPA ecosystem should not be neglected [@Martinez2021].

@@

The MPA N-BB I presents complex oceanographic conditions that generate an internal spatial heterogeneity, mainly along its longitudinal axis [@Matano2019]. So far this heterogeneity has been reflected in phytoplankton and zooplankton communities [@Bertola2018; @GarciaAlonso2020; @Spinelli2020], and in fish assemblages [@Delpiani2020]. Moreover, seasonal variations also occur in some physical and biological aspects of the MPA N-BB I [@GarciaAlonso2018; @Matano2019]. Considering both MPAs (N-BB I and II), a seasonal variation in the community composition of marine mammals and seabirds was recorded recently [@Dellabianca2023]. The spatial and seasonal variations in the plankton community might affect the energy and matter flow to higher levels of the food web. This has been recently studied in the vicinity of the MPA N-BB I, in the Beagle Channel, where a differential energy flow pattern of the plankton community has been recognised in two micro-basins of the Channel separated by a sill, each with different physicochemical properties [@Giesecke2021], nutrient concentration [@Latorre2023] as well as in the dominant component of the plankton community [@Bruno2023a; @Presta2023]. Although we were aware of the above, we decided to characterise a food web representing the whole MPA N-BB I year round since this is the first study of its type in the area.

@@

We identified several species as important regarding different aspects of trophic structure and functioning, and response to perturbations (i.e. environmental/anthropogenic changes). On the one hand, we suggest that generalist species, mainly fishes, play a crucial role in the ecosystem's bentho-pelagic coupling process. At the same time, we propose that other species besides the longtail southern cod *Patagonotothen ramsayi* and the Fueguian sprat *Sprattus fuegensis* should be considered relevant energy transfers for the ecosystem. Finally, we argue that it is the diversity of species, representing the benthic and pelagic habitats, that maintains the connectivity of the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

Our work provides valuable information for supporting the management of the MPA N-BB in the sense that it disentangles the multiple ecosystem players and their roles. Moreover, by building the network of predator-prey interactions we have set the baseline to model the environmental and anthropogenic perturbations affecting the MPA ecosystem. This knowledge becomes of importance at the regional scale given that the MPA ecosystem is a known spawning ground for many species of commercial interest [@AdministraciondeParquesNacionales2022] and a potential area of high carbon storage and sequestration due to the characteristics of the benthic community [@Bax2022; @Bergagnainrev].

@@

```{r figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 6.5, fig.align='center', fig.cap="Species-level properties by trophic level: (A) betweenness, (B) closeness, (C) trophic similarity, and (D) topological role. Each point represents a species. Note that for panels A, B and C only species with TLs equal or greater than 2 were considered. In the boxplot (D), the horizontal line in the middle of each box represents the median of the data, the colored box represents the interquartile range (range between the 1st, or 25th percentile, and 3rd, or 75th percentile, quartile), containing the middle 50% of data points."}
# Trophic similarity, Closeness, Betweenness & Topological role by Trophic level

>>>>

---
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
    keep_tex: yes
bibliography: AMPBancoBurdwood.bib
csl: "elsevier-harvard2.csl"
documentclass: article
geometry:
- vmargin=1in
- hmargin=1in
header-includes:
- \usepackage{amsmath}
- \usepackage{pdflscape,booktabs}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage[running]{lineno}
- \linenumbers
---

@@

The world’s oceans designated under marine protection have increased recently. Most Marine Protected Areas (MPAs) target vulnerable, keystone, charismatic and or endemic species. In the sub-Antarctic, ocean protection is associated to oceanic islands, except for MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, ~53º–55ºS and ~56º–62ºW), which are associated to a submarine plateau and a southern deep slope, respectively. Here, we present the first analysis of the predator-prey network for the MPA N-BB. We applied a topological network approach to characterise the complexity and structure of the food web, and identify the species’ role in such a framework. The MPA N-BB food web consisted of 1788 interactions and 379 species, with a connectance of 0.01. Almost half of the consumers feed at more than one trophic level (0.48), and the network displayed a small-world pattern (short path length, high clustering of compartments). This network pattern suggests that the ecosystem might be vulnerable to perturbations targeting highly connected species, although some properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Several species arose as important for the trophic structure and functioning, and response to perturbations. Generalist species, mainly fishes, play a crucial role in the benthopelagic coupling and should be considered as relevant energy transfers for the ecosystem. We argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for securing the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

@@

Maintaining ocean ecosystem services relies on a healthy ocean, and biological species play a key role in this process. It is essential to identify the main species involved in sustaining this health for effective management and conservation purposes. In this regard, @Bergagna2020 have identified the principal ecosystem services that are provided by species inhabiting the MPA N-BB; for instance, nursery and feeding areas for commercially important fishes (e.g. Patagonian grenadier *Macruronus magellanicus*, Patagonian toothfish *Dissostichus eleginoides*), water purification by filter feeders (e.g. sponges), among others. These species are part of a complex system in terms of biodiversity. There is robust knowledge on the complexity considering the richness of the benthic and plankton communities in the MPA N-BB ecosystem [@Schejter2016; @Guinder2020; @Schejter2020; @AdministraciondeParquesNacionales2022]. Overall, 811 benthic and plankton species have been identified in the MPA N-BB ecosystem, where 349 were reported for the first time in the area in recent years [@AdministraciondeParquesNacionales2022]. Recently, the structure of the southwestern South Atlantic Ocean has been hypothesised to be under a 'wasp-waist' control, meaning that the structure and dynamics of the ecosystem are regulated primarily by mid-trophic level species (e.g., fishes, crustaceans) [@Padovani2012; @Saporiti2015; @Riccialdelli2020]. In particular, the ecosystem of the MPA N-BB shows a more pronounced 'wasp-waist' structure, meaning a shorter average food chain and a greater trophic overlap and redundancy, respect to other sub-Antarctic areas such as the continental shelf off Tierra del Fuego. The high abundance of a limited number of prey species at the intermediate trophic level of the food web which leads to a specific feeding reliance among high-trophic level predators. This phenomenon shortens food-web length while simultaneously amplifying trophic overlap and redundancy within the overall food web. The Fuegian sprat *Sprattus fuegensis* and longtail southern cod *Patagonotothen ramsayi* has been considered the most plausible 'wasp-waist' species for the MPA N-BB [@Riccialdelli2020], since they are species with high regional abundances [@Madirolas2000; @Arkhipkin2013], they occupy mid-trophic levels [@Riccialdelli2010; @Riccialdelli2020], many predators feed on them [@Riccialdelli2020], and their population dynamics appear to depend on the environment (e.g. climatic variations) [@Diez2018]. These recent efforts have increased our understanding of the biological communities and their complexity, but an understanding of species interactions’ complexity and structure is lacking. Such understanding will provide a more complete picture of the ecosystem, generate a baseline to comprehend climate change effects [@Franco2020] and serve as an input to decision makers on conservation actions [@AdministraciondeParquesNacionales2022].

This aspect can be tackled by analysing one of the most frequent relationships between species: the predator-prey interaction [@Bascompte2009]. The assembly of these interactions in a particular region is referred to as a food web, representing the roadmap for matter and energy flow in an ecosystem. In recent years, topological network approaches have been successfully applied to study complex high-latitude marine ecosystems, improving our knowledge on structure, functioning, and response to environmental/anthropogenic changes [@Kortsch2015; @Cordone2018; @Funes2022; @Marina2023]. Among anthropogenic threats, it is worth mentioning that contaminants like mercury and microplastics have been recently reported as important threats to the MPA N-BB region [@Cossi2021; @DiMauro2022; @Fioramonti2022]; also fishing vessels are allowed to operate in the western section of the MPA N-BB (i.e. Marine National Reserve category), altering the stocks of commercially important fish species [@Martinez2021; @AdministraciondeParquesNacionales2022]. Moreover, there is a potential hazard related to the effects of offshore activities (exploration and exploitation of gas and oil) to the west of the MPA N-BB [@AdministraciondeParquesNacionales2022].

@@

where $E_i$ is the effective number of interactions between $K_i$ nearest-neighbour nodes of node $i$ and the maximal possible number of such interactions [@Newman2003].
To test whether the food web presented the small-world pattern, we compared the empirical values of $CPL$ and $CC$ with those resulting from 1000 randomly generated networks with the same size ($S$) and number of interactions ($L$). Random webs were created using the Erdös-Rényi model, where links are added to the complete set of nodes ($S$) and chosen uniformly randomly from the set of all possible links [@Erdos1959]. The preciseness of our method lies in the use of confidence intervals (CI 99%) for the empirical-random comparison of the $CPL$ and $CC$ properties. If the empirical value is positioned within or to the left (= lower than) the CI 99% of the random CPL, and to the right (= higher than) the CI 99% of the CC, then the food web is considered to present the SW topology [@Marina2018a].

@@

The species-level analysis aims to describe the species' role in the food web. For this, we considered the following properties: betweenness $Btw$, closeness $Cl$, trophic similarity $TS$, topological role $TR$, and trophic level $TL$ (Table 1). Topological roles refer to the fact that food webs tend to naturally organise in non-random, modular patterns, where modules are defined as a group of species that interact more frequently among themselves than with species that are not members of the module [@Guimera2005]. This means that groups of prey and predators interact more strongly with each other, than with species belonging to other groups. Consequently, modularity measures how strongly these modules interact with each other compared to their interaction with species from other modules. Species can play different roles in this respect, according to the pattern of interactions within their own module and/or across modules. We computed the topological role for each species, classified as “module hub”, species with a relatively high number of interactions, but most within its own module; “module specialist”, species with relatively few interactions and most within its own module; “module connector”, species with relatively few interactions mainly between modules; and “network connector”, species with high connectivity between and within modules [@Guimera2005]. Refer to the Supplementary Material for the equations used to calculate the species-level properties.

@@

The degree distribution of the food web showed an asymmetric frequency in the number of interactions, where most of the species had a relatively low number of interactions and few species concentrated most of them (Figure 3A). The distribution of prey among predators showed that most consumers fed on a low number of prey whereas few had multiple prey (Figure 3B). The top-five predators in number of prey were: yellowfin notothen *Patagonotothen guntheri* (Notothenioid fish, 50 prey), rock cod *Patagonotothen ramsayi* (Notothenioid fish, 49 prey), broad nose skate *Bathyraja brachyurops* (Chondrichthyan, 33 prey), Patagonian toothfish *Dissostichus eleginoides* (Notothenioid fish, 30 prey), and graytail skate *Bathyraja griseocauda* (Chondrichthyan, 28 prey). Following the same distribution pattern, few prey were targeted by multiple predators (Figure 3C). The top-five prey (or food sources) in number of predators were: Detritus (Non-living, 153 predators), the three categories of Diatoms considered (benthic, centric and pennate, 72.5 predators on average), and species of the genus *Euphausia* (Zooplankton, 46 predators). Finally, aggregating the interactions by functional group (Figure 3D), most interactions were concentrated in a few species. The most evident species were: *Doryteuthis gahi* (Cephalopoda), *Grimothea (=Munida) gregaria* (Decapoda), *Patagonotothen ramsayi*, *Patagonotothen guntheri* and *Dissostichus eleginoides* (bentho-pelagic fish), *Sprattus fuegensis* and *Micromesistius australis* (pelagic fish), and species of *Euphausia* and *Themisto gaudichaudii* (Zooplankton).
Overall, there is an evident asymmetry in the distribution of interactions among species at different levels (i.e. considering the entire food web, gathered by prey, predator and functional group) in the MPA N-BB food web. A list of the distribution of interactions per species is presented in Supplementary Material (Table S3).

@@

Food web connectance is a feature that summarises the complexity of the network, but more importantly, it is an emergent property of pairwise species interactions [@Poisot2014]. It contains information regarding how interactions within an ecological network are distributed and predicts reasonably well key dynamical properties of ecological networks [@Dunne2002a]. Complex marine food webs (i.e. with more than 25 trophic species) show connectance values ranging from 0.01 - 0.27 [@Marina2018a]. In particular, food webs from high-latitude regions tend to exhibit a connectance closer to the minimum (between 0.01 and 0.05) [@Santana2013; @Kortsch2015; @Rodriguez2022]. The connectance of the food web of the MPA N-BB (0.01) is one of the lowest reported so far for these regions; in particular, it appears to be much lower than that of Beagle Channel (0.05), an adjacent coastal area [@Rodriguez2022]. Whether food webs display a low or a high connectance helps to better comprehend ecosystem's synthetic properties like robustness. Empirical research indicates that ecological networks with high connectance are resilient to external perturbations, including the introduction of new species like invasive ones [@Smith-Ramesh2017], as well as the removal of species due to local extinction [@Dunne2002; @Montoya2003].

@@

Closeness and betweenness are defined as "mesoscale" properties because they consider direct and indirect interactions, therefore describing the focal species’ ability to influence the rest of the species of the food web [@Freeman1978; @Lai2012].
Closeness quantifies how many steps away species $i$ is from all other species in the food web, and is proportional to how rapidly the indirect effects of the focal species can spread to other species in the network [@Scotti2010].
Considering the relationship between this property and TL in the food web of the MPA N-BB, it is worth highlighting that species with the highest closeness values are low-TL consumers (TL = 2 - 3): species of the zooplankton community, *Calanus* and *Euphausia*, *Zygochlamys patagonica* (Bivalvia), and Brachiopoda. Any perturbation affecting these species, such as the recently confirmed contaminants mercury [@Fioramonti2022] and microplastics [@Cossi2021; @DiMauro2022], and the warming trend reported for the surface and subsurface waters in the Southwest Atlantic Ocean [@Franco2020a; @Galvan2022a], should be of concern since it might reach many other species in the food web.
Otherwise, betweenness measures the number of shortest paths between species, providing information on the importance of species as “bridges” for energy transfer: a species with high betweenness takes part in more food chains and therefore affects more energy flows [@Scotti2010]. We have identified the longtail southern cod *Patagonotothen ramsayi* as the most important species in this sense. Moreover, in light of our analysis, other species like the Patagonian toothfish *Dissostichus eleginoides* (juveniles), the Patagonian cod *Salilota australis*, the yellowfin notothenioid *Patagonotothen guntheri*, and the Patagonian longfin squid *Doryteuthis gahi* should be considered as relevant in the energy transfer in the ecosystem. All these species have a mid-trophic position in the food web, supporting the 'wasp-waist' control hypothesis for the MPA N-BB [@Riccialdelli2020].

Ecosystems with a pronounced 'wasp-waist' structure are suggested to present a high trophic redundancy, since many species would show similar trophic habits [@Cury2000]. Such trophic redundancy is expected to be greater for species in TLs above the waist that depend on a very few species at this intermediate level. However, our results show a negative relationship between trophic similarity (or redundancy) and trophic level, indicating functional similarity at low and mid-TL species compared to higher TL species for the MPA N-BB food web. This contradiction might be reflecting the potential discrepancy that could arise between qualitative (this study) and quantitative food web studies. Here it is important to note that our analyses considered all potential predator-prey interactions in the food web, assigning equal importance to each of them. Although this could be a disadvantage when inferring on energy flows, one of the main advantages of qualitative food web studies is that it considers the diversity of the communities inhabiting a particular study area, therefore providing a more representative picture of all components (species) acting on the ecosystem. In this sense, our study highlights the uniqueness in terms of the trophic role of high-TL predators. Not only the expected pelagic animals such as marine mammals and seabirds arise as relevant, but also demersal vertebrate (chondrichthyans *Bathyraja macloviana* and *Squalus acanthias*) and benthic invertebrate species (echinoderms *Diplopteraster clarki* and *Pteraster sp.*) are noteworthy. The role that such species play in the MPA N-BB ecosystem is unique and perturbations on them might result in unprecedented changes at the trophic structure and functioning level. In this regard, we should mention the potential threat of the fisheries operating in the western section of the MPA N-BB, where this activity is allowed and mostly focuses on the Patagonian toothfish *Dissostichus eleginoides* and the southern blue whiting *Micromesistius australis* [@Martinez2015]. Although the fishing effort is concentrated outside the limits of the MPA N-BB, the impact on the MPA ecosystem should not be neglected [@Martinez2021].

@@

The MPA N-BB I presents complex oceanographic conditions that generate an internal spatial heterogeneity, mainly along its longitudinal axis [@Matano2019]. So far this heterogeneity has been reflected in phytoplankton and zooplankton communities [@Bertola2018; @GarciaAlonso2020; @Spinelli2020], and in fish assemblages [@Delpiani2020]. Moreover, seasonal variations also occur in some physical and biological aspects of the MPA N-BB I [@GarciaAlonso2018; @Matano2019]. Considering both MPAs (N-BB I and II), a seasonal variation in the community composition of marine mammals and seabirds was recorded recently [@Dellabianca2023]. The spatial and seasonal variations in the plankton community might affect the energy and matter flow to higher levels of the food web. This has been recently studied in the vicinity of the MPA N-BB I, in the Beagle Channel, where a differential energy flow pattern of the plankton community has been recognised in two micro-basins of the Channel separated by a sill, each with different physicochemical properties [@Giesecke2021], nutrient concentration [@Latorre2023] as well as in the dominant component of the plankton community [@Bruno2023a; @Presta2023]. Although we were aware of the above, we decided to characterise a food web representing the whole MPA N-BB I year round since this is the first study of its type in the area. In future studies, environmental variables such as depth, current speed and/or nutrient availability should be included in food web works, as done recently [@Cordone2020].

@@

We identified several species as important regarding different aspects of trophic structure and functioning, and response to perturbations (i.e. environmental/anthropogenic changes). On the one hand, we suggest that generalist species, mainly fishes, play a crucial role in the ecosystem's bentho-pelagic coupling process. At the same time, we propose that other species besides the longtail southern cod *Patagonotothen ramsayi* and the Fuegian sprat *Sprattus fuegensis* should be considered relevant energy transfers for the ecosystem. Finally, we argue that it is the diversity of species, representing the benthic and pelagic habitats, that maintains the connectivity of the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

Our work provides valuable information for supporting the management of the MPA N-BB in the sense that it disentangles the multiple ecosystem players and their roles. Moreover, by building the network of predator-prey interactions we have set the baseline to model the environmental and anthropogenic perturbations affecting this oceanic ecosystem. This knowledge becomes of importance at the regional scale given that the MPA N-BB ecosystem is a known spawning ground for many species of commercial interest [@AdministraciondeParquesNacionales2022] and a potential area of high carbon storage and sequestration due to the characteristics of the benthic community [@Bax2022; @Bergagnainrev].

@@

```{r figure4, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 6.5, fig.height = 6.5, fig.align='center', fig.cap="Species-level properties by trophic level: (A) betweenness, (B) closeness, (C) trophic similarity, and (D) topological role. Each point represents a species. Note that for panels A, B and C only species with TLs equal or greater than 2 were considered. In the boxplot (D), the horizontal line in the middle of each box represents the median of the data, the coloured box represents the interquartile range (range between the 1st, or 25th percentile, and 3rd, or 75th percentile, quartile), containing the middle 50% of data points."}
# Trophic similarity, Closeness, Betweenness & Topological role by Trophic level
